### 2018-11-19 ms-meetup-nov18.Rmd
initial commit

---
title: "Using R to help plan the <br> future of transport"
subtitle: "Münster R Users Group"
author: "Mark Padgham <br> Active Transport Futures"
date: "Tuesday 20th November, 2018 "
output:
      xaringan::moon_reader:
            yolo:
                        img: img/fortune-beach.png
                        times: 2
            lib_dir: libs
            css: xaringan-themer.css
            nature:
                  highlightStyle: github
                  highlightLines: true
                  countIncrementalSlides: false
---
```{r load, include = FALSE, eval = TRUE}
library(xaringanthemer)
source ("myxaringantheme.R")
source ("script.R")
promenade <- readRDS ("data/osm-promenade.Rds")
ms48155 <- readRDS ("data/osm-ms-48155.Rds")
net <- f <- readRDS ("data/dodgr-flows-ms.Rds")
net$flow <- NULL
```

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)# suppress version num in subdir name

knitr::opts_chunk$set(cache = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

class: left, middle, inverse

.left-column[
`r icon::fa_github(size = 2, animate = "spin")` atfutures

`r icon::fa_github(size = 2, animate = "spin")` mpadge

`r icon::fa_github(size = 2, animate = "spin")` ropensci
]

.right-column[
`r icon::fa_twitter(size = 2, animate = "spin")` bikesRdata

`r icon::fa_envelope(size = 1)` .small[mark.padgham@email.com]<br><br>

`r icon::ii_ios_world_outline(size = 5)` atfutures.github.io
]

.box-bottom[
slides at <br>
[https://github.com/mpadge/ms-user-meetup-nov2018](https://github.com/mpadge/ms-user-meetup-nov2018)
]

---
class: left, top, inverse

### Primary Collaborators

- Robin Lovelace (@robinlovelace)
- Layik Hama (@layik)
- Andreas Petutschnig (@karpfen)
- Andrew Smith (@virgesmith)

### Additional `r icon::fa_github(animate = "spin")` Contributions From ...

.left-column[
- RichardEllison
- tbuckl
- skcott
- hrbrmstr
- maelle
- mdsumner
- mem48
- jimshady
- angela-li
  ]

.right-column[
- graceli8
- arfon
- jimhester
- tazinho
- karthik
- fzenoni
- mhenderson
- patperu
- eyesofbambi
  ]

---
class: left, top, inverse

# Active Transport Futures

---
background-image: url(img/amsterdam-cars.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000
class: left, top, inverse

# Active Transport Futures

---
class: left, top, inverse
background-image: url(img/ms-promenade.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000

# Active Transport Futures

---
class: left, top, inverse
background-image: url(img/kl-bike-path.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000

# [Active Transport Futures](http://www.who.int/sustainable-development/cities/en/)

---

# Active Transport Futures

For a given city:

- Collate data on patterns of urban mobility
- Collate data on transport networks <br> (road, pedestrian, public transport)
- Analyse "flows" throughout the network(s)
- Collate data on (spatial patterns of) air pollution
- Analyse dynamic interactions between movement, air pollution, and health
- Develop, compare, contrast scenarios for ameliorating negative effects while
  enhancing positive effects
- [Health Economic Assessment Tool (HEAT)](https://heatwalkingcycling.org/#homepage)

---

# Active Transport Futures

For a given city:

- ~~Collate data on patterns of urban mobility~~
- Collate data on transport networks <br> (road, pedestrian, public transport)
- Analyse "flows" throughout the network(s)
- ~~Collate data on (spatial patterns of) air pollution~~
- Analyse dynamic interactions between movement, air pollution~~, and health~~
- ~~Develop, compare, contrast scenarios for ameliorating negative effects while
  enhancing positive effects~~
- ~~[Health Economic Assessment Tool (HEAT)](https://heatwalkingcycling.org/#homepage)~~

---

# Active Transport Futures

## Software: R Packages

- `osmdata` (Open Street Map data)
- `dodgr` (Distances On Directed GRaphs)
- `stplanr` (Sustainable Transport PLANning in R)
- `osmplotr` (OSM plotting routines)
- `geoplumber` (Turbo-charged web server for geo data)
- `bikedata` (Data from public hire bicycles)
- `spatialcluster`
- ... plus more

- demo of `geoplumber` at [`http://35.233.61.182`](http://35.233.61.182)

---

# Active Transport Futures

## Software: R Packages

- `osmdata` (Open Street Map data)
- `dodgr` (Distances On Directed GRaphs)
- ~~`stplanr` (Sustainable Transport PLANning in R)~~
- ~~`osmplotr` (OSM plotting routines)~~
- ~~`geoplumber` (Turbo-charged web server for geo data)~~
- ~~`bikedata` (Data from public hire bicycles)~~
- ~~`spatialcluster`~~
- ... plus more

- demo of `geoplumber` at [`http://35.233.61.182`](http://35.233.61.182)

---
class: left, top

# The osmdata package

OSM = Open Street Map

- CRAN
- `r icon::fa_github(size = 1, animate = "spin")` `ropensci/osmdata`
- Several mentions on
  [OSM wiki](https://wiki.openstreetmap.org/wiki/OSM_Scientific_Tools)

--

## Why OSM data?

---
class: center, top, inverse
.fontBIG[
Popquiz: <br><br>What is the least mapped country on the planet?
]

---
class: center, middle, inverse
background-image: url(img/osm-north-korea.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000

---
class: center, middle, inverse
background-image: url(img/anju-bing.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000

---
class: center, middle, inverse
background-image: url(img/anju-google.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000

---
class: center, middle, inverse
background-image: url(img/anju-google2.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000

---
class: center, middle, inverse
background-image: url(img/anju-osm.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000

---

# The osmdata package

- CRAN
- `r icon::fa_github(size = 1, animate = "spin")` `ropensci/osmdata`
- Several mentions on
  [OSM wiki](https://wiki.openstreetmap.org/wiki/OSM_Scientific_Tools)

## Why OSM data?

--

Because Open Street Map is the most comprehensive street atlas humanity has ever
created.

## How?

`osmdata` uses the `overpass` server to retrieve data

---

# The osmdata package

```{r opq1}
library (osmdata)
opq ("muenster de")
```

---

# The osmdata package

```{r opq2}
library (osmdata)
opq ("48155 de")
```

---

# The osmdata package

```{r add-osm-ft}
library (osmdata)
opq ("48155 de") %>%
    add_osm_feature (key = "highway")
```

---

# The osmdata package

```{r osmdata-sf, eval = FALSE}
library (osmdata)
opq ("48155 de") %>%
    add_osm_feature (key = "highway") %>%
    osmdata_sf ()
```
```{r osmdata-sf-fakey, eval = TRUE, echo = FALSE, message = TRUE}
print (ms48155)
```

---

# The osmdata package

```{r osmdata-sf2, eval = FALSE}
library (osmdata)
opq ("48155 de") %>%
    add_osm_feature (key = "highway") %>%
    osmdata_sf ()
```

- `opq` `r icon::fa_long_arrow_alt_right()` .bold[WHERE]
- `opq` `r icon::fa_long_arrow_alt_right()` .bold[WHERE]
- `add_osm_feature` `r icon::fa_long_arrow_alt_right()` .bold[WHAT]
- `osmdata_sf` `r icon::fa_long_arrow_alt_right()` .bold[HOW]

---

# The osmdata package - Spatial Formats

```{r osmdata-sf3, eval = FALSE}
library (osmdata)
opq ("48155 de") %>%
    add_osm_feature (key = "highway") %>%
    osmdata_sf ()
```

- `osmdata_sf` `r icon::fa_long_arrow_alt_right()` .bold[Simple Features]

---

# The osmdata package - Spatial Formats

```{r osmdata-sp, eval = FALSE}
library (osmdata)
opq ("48155 de") %>%
    add_osm_feature (key = "highway") %>%
    osmdata_sp ()
```

- `osmdata_sf` `r icon::fa_long_arrow_alt_right()` .bold[Simple Features]
- `osmdata_sp` `r icon::fa_long_arrow_alt_right()` .bold[R Spatial]

---

# The osmdata package - Spatial Formats

```{r osmdata-sc, eval = FALSE}
library (osmdata)
opq ("48155 de") %>%
    add_osm_feature (key = "highway") %>%
    osmdata_sc ()
```

- `osmdata_sf` `r icon::fa_long_arrow_alt_right()` .bold[Simple Features]
- `osmdata_sp` `r icon::fa_long_arrow_alt_right()` .bold[R Spatial]
- `osmdata_sc` `r icon::fa_long_arrow_alt_right()` .bold[Silicate (Simplicial Complex)]

---

# The osmdata package - Spatial Formats

```{r osmdata-xml, eval = FALSE}
library (osmdata)
opq ("48155 de") %>%
    add_osm_feature (key = "highway") %>%
    osmdata_xml (file = "ms48155.osm")
```

- `osmdata_sf` `r icon::fa_long_arrow_alt_right()` .bold[Simple Features]
- `osmdata_sp` `r icon::fa_long_arrow_alt_right()` .bold[R Spatial]
- `osmdata_sc` `r icon::fa_long_arrow_alt_right()` .bold[Silicate (Simplicial Complex)]
- `osmdata_xml` `r icon::fa_long_arrow_alt_right()` .bold[XML / OSM]

---

# The osmdata package - Spatial Formats

```{r osmdata-pbf, eval = FALSE}
library (osmdata)
opq ("48155 de") %>%
    add_osm_feature (key = "highway") %>%
    osmdata_pbf (file = "ms48155.pbf")
```

- `osmdata_sf` `r icon::fa_long_arrow_alt_right()` .bold[Simple Features]
- `osmdata_sp` `r icon::fa_long_arrow_alt_right()` .bold[R Spatial]
- `osmdata_sc` `r icon::fa_long_arrow_alt_right()` .bold[Silicate (Simplicial Complex)]
- `osmdata_xml` `r icon::fa_long_arrow_alt_right()` .bold[XML / OSM]
- `osmdata_pbf` `r icon::fa_long_arrow_alt_right()` .bold[PBF (Protocol Buffer Format)]

---

# The osmdata package

```{r osmdata_sf2, eval = FALSE, message = TRUE}
library (osmdata)
ms48155 <- opq ("48155 de") %>%
    add_osm_feature (key = "highway") %>%
    osmdata_sf ()
ms48155
```
```{r osmdata_sf2-fakey, echo = FALSE, message = TRUE}
print (ms48155)
```
---

# The osmdata package - Visualisation

```{r mapview, eval = FALSE}
library (osmdata)
ms48155 <- opq ("48155 de") %>%
    add_osm_feature (key = "highway") %>%
    osmdata_sf ()
library (mapview)
mapview (ms48155$osm_lines)
```

---
background-image: url(img/ms48155lines-mapview.png)
background-size: contain 
background-position: 50% 50%
class: center, top, inverse

---

# The osmdata package - Visualisation

How to visualise just the Promenade?

```{r mapview2, eval = FALSE}
library (osmdata)
ms48155 <- opq ("48155 de") %>% #<<
    add_osm_feature (key = "highway") %>%
    osmdata_sf ()
```
---

# The osmdata package - Visualisation

How to visualise just the Promenade?

```{r promenade-data1, eval = FALSE, message = TRUE}
library (osmdata)
promenade <- opq ("münster de") %>% #<<
    add_osm_feature (key = "highway") %>%
    osmdata_sf ()
```

---

# The osmdata package - Visualisation

How to visualise just the Promenade?

```{r promenade-data2, eval = FALSE}
library (osmdata)
promenade <- opq ("münster de") %>%
    add_osm_feature (key = "highway") %>%
    add_osm_feature (key = "name", value = "promenade", #<<
                     value_exact = FALSE) %>%           #<<
    osmdata_sf ()
```

---

# The osmdata package - Visualisation

How to visualise just the Promenade?

```{r promenade-data3, eval = FALSE}
library (osmdata)
promenade <- opq ("münster de") %>%
    add_osm_feature (key = "highway") %>%
    add_osm_feature (key = "name", value = "Promenade",
                     value_exact = FALSE) %>%
    osmdata_sf ()
library (mapview)
mapview (promenade$osm_lines) #<<
```

---
background-image: url(img/mspromenade-mapview.png)
background-size: contain 
background-position: 50% 50%
class: center, middle, inverse

---

class: middle, inverse, center

.fonthuge[
Digression:<br>Visualising OSM data
]

---

# Visualising osmdata

```{r prom-sf, message = TRUE}
promenade
```

---

# Visualising osmdata

```{r prom-sf-print}
library (sf)
promenade$osm_lines
```
---

# Visualising osmdata

```{r, eval = FALSE}
library (mapview)
mapview (ms48155$osm_lines)
```

---
background-image: url(img/ms48155lines-mapview.png)
background-size: contain 
background-position: 50% 50%
class: center, middle, inverse

---

# Visualising osmdata

```{r, eval = FALSE}
library (mapview)
mapview (ms48155$osm_lines) %>%
    addFeatures (promenade$osm_lines)
```
... currently will not work (coz of `leaflet`; it's not `mapview`'s fault)

--

- but [`deck.gl`](https://deck.gl) will

--

- and it's available in R via the `mapdeck` package brought to us by 
`r icon::fa_github(animate = "spin")` `symbolixAU`

---

# Visualising osmdata

```{r promenade-data4, eval = FALSE}
promenade <- opq ("münster de") %>%
    add_osm_feature (key = "highway") %>%
    add_osm_feature (key = "name", value = "Promenade",
                     value_exact = FALSE) %>%
    osmdata_sf ()
```

---

# Visualising osmdata

A little bit of pre-processing #1: Because `mapdeck` is highly experimental and
currently needs to know where to centre the map.

```{r mapdeck-pre}
location <- ms48155$osm_lines %>%
    sf::st_coordinates () %>%
    apply (2, mean) %>%
    as.numeric () %>%
    magrittr::extract (1:2)
location
```

---

# Visualising osmdata

A little bit of pre-processing #2: Because `mapdeck` needs a personal mapbox
token which can be set with `set_token()`, and automatically used to generated
all maps

```{r mapdeck-token, eval = FALSE}
library (mapdeck)
set_token (<my_mapbox_token>)
```

---

# Visualising osmdata

```{r mapdeck3, eval = FALSE, echo = TRUE}
library (mapdeck)
mapdeck (style = 'mapbox://styles/mapbox/dark-v9',
         location = location,
         zoom = 12) %>%
    add_path (data = ms48155$osm_lines,
              layer_id = "ms",
              stroke_width = 1,
              stroke_colour = "#4444dd",
              stroke_opacity = 80) %>%
    add_path (data = promenade$osm_lines,
              layer_id = "promenade",
              stroke_width = 5,
              stroke_colour = "#33cc33",
              stroke_opacity = 100)
```
---
class: center, middle, inverse
background-image: url(img/ms-highway-promenade-mapview.png)
background-size: contain 
background-position: 50% 50%

---

# Visualising osmdata

One further alternative for static plots is the [`osmplotr`
package](https://github.com/ropensci/osmplotr)

---
class: top, center, inverse

.fontBIG[
Question:<br><br> Has anybody <br>used `osmdata`?
]

---
class: top, center, inverse

.fontBIG[
Question:<br><br> What kind of default plot method might <br>be useful?
]

---

# osmdata

For more information, see [the website](https://github.com/ropensci/osmdata)

---
class: top, center, inverse

.fonthuge[
    **dodgr**<br>
    ]
.fontBIG[
    Distances on Directed GRaphs
]

---
class: top, center, inverse

.fonthuge[
    **dodgr**<br>
    ]

* CRAN <br><br>
* `r icon::fa_github(size = 4, animate = "spin", colour = "#ffcc00")` [`ATFutures/dodgr`](https://github.com/atfutures/dodgr)

---

# dodgr

Get all OSM highway data for Münster (not just 48155)

```{r osm-ms, eval = FALSE}
ms <- opq ("münster de") %>%
    add_osm_feature (key = "highway") %>%
    osmdata_sf ()
ms
```
```{r load-ms, echo = FALSE, eval = FALSE, message = TRUE}
ms <- readRDS ("data/osm-ms-highways.Rds")
ms
```
```{r load-ms-fakey, echo = FALSE, message = TRUE}
message ("Object of class 'osmdata' with:\n",
"                 $bbox : 51.8401448,7.4737853,52.0600251,7.7743634\n",
"        $overpass_call : The call submitted to the overpass API\n",
"                 $meta : metadata including timestamp and version numbers\n",
"           $osm_points : 'sf' Simple Features Collection with 157614 points\n",
"            $osm_lines : 'sf' Simple Features Collection with 33545 linestrings\n",
"         $osm_polygons : 'sf' Simple Features Collection with 510 polygons\n",
"       $osm_multilines : 'sf' Simple Features Collection with 0 multilinestrings\n",
"    $osm_multipolygons : 'sf' Simple Features Collection with 18 multipolygons")
```

---

# dodgr

Convert Münster OSM data to `dodgr` format

```{r weight-streetnet-ms1, eval = FALSE, message = TRUE}
ms <- osm_poly2line (ms) # osmdata function
```

---

# dodgr

Convert Münster OSM data to `dodgr` format

```{r weight-streetnet-ms2, eval = FALSE, message = TRUE}
ms <- osm_poly2line (ms) # osmdata function
library (dodgr)
net <- weight_streetnet (ms$osm_lines, wt_profile = "bicycle") #<<
```

---

# dodgr

Convert Münster OSM data to `dodgr` format

```{r weight-streetnet-ms3, eval = FALSE, message = TRUE}
ms <- osm_poly2line (ms) # osmdata function
library (dodgr)
net <- weight_streetnet (ms$osm_lines, wt_profile = "bicycle") #<<
dim (net); head (net)
```
```{r weight-streetnet-ms-fakey, eval = TRUE, echo = FALSE, message = TRUE}
library (dodgr)
dim (net); head (net)
```

---

# dodgr distances

```{r ms-components, eval = TRUE, message = TRUE}
dim (net)
head (table (net$component), 10)
```

---

# dodgr distances

Calculate cycling distances between random points.

```{r dodgr-vertices1-fakey, eval = FALSE}
v <- dodgr_vertices (net) #<<
dim (v); head (v)
```
```{r dodgr-vertices1, eval = TRUE, echo = FALSE}
v <- dodgr_vertices (net)
dim (v); head (v, n = 5)
```

- `id` is the unique OSM is of each vertex<br>("node" in OSM terminology)
- most `dodgr` functions work either with<br>coordinates or `id` values

---

# dodgr distances

Calculate cycling distances between random points.

```{r dodgr-vertices2, eval = FALSE}
v <- dodgr_vertices (net)
dim (v); head (v)
```
```{r dodgr-vertices2-fakey, eval = TRUE, echo = FALSE, message = TRUE}
dim (v); head (v, n = 5)
```
```{r vert-sample, eval = TRUE}
pts <- sample (v$id, size = 1000)
head (pts)
```

---

# dodgr distances

```{r dodgr-dists-help-fakey, echo = TRUE, eval = FALSE}
?dodgr_dists
```
```{r dodgr-dists-help, echo = FALSE, message = TRUE}
helpfile <- utils:::.getHelpFile (help ("dodgr_dists"))
hs <- capture.output (tools:::Rd2txt(helpfile))
#hs <- capture.output (tools:::prepare_Rd(helpfile))
cat(hs, sep="\n")
```

---

# dodgr distances

Calculate cycling distances between random points.

```{r dodgr-vertices3, eval = FALSE}
v <- dodgr_vertices (net)
dim (v); head (v)
```
```{r dodgr-vertices3-fakey, eval = TRUE, echo = FALSE, message = TRUE}
dim (v); head (v, n = 5)
```
```{r vert-sample3, eval = TRUE, message = TRUE}
pts <- sample (v$id, size = 1000)
head (pts)
```

---

# dodgr distances

```{r dodgr-dists1, eval = FALSE}
system.time (
             d <- dodgr_dists (net, from = pts, to = pts) #<<
             )
```
```{r dodgr-dists1-fakey, echo = FALSE, message = TRUE}
message ("   user  system elapsed\n",
" 43.847   0.013   6.042")
```

--

- Those 1,000,000 distances could have been calculated with the google API
- And that would cost `1e6 * 0.004 = $US4,000`
- $US 4,000 for 6 seconds of calculation!
- To date, `dodgr` provides one of the only .bold[free] ways to perform such
  calcualtions, and is also entirely .bold[open source].

---

# dodgr distances

```{r dodgr-dists1-copy, eval = FALSE}
system.time (
             d <- dodgr_dists (net, from = pts, to = pts)
             )
```
```{r dodgr-dists1-copy-fakey, echo = FALSE, message = TRUE}
message ("   user  system elapsed\n",
" 43.847   0.013   6.042")
```
```{r dodgr-dists-result1, eval = FALSE}
class (d); dim (d); range (d); range (d, na.rm = TRUE)
```
```{r dodgr-dists-result1-fakey, eval = TRUE, echo = FALSE, message = TRUE}
message ("[1] matrix\n",    # explicit print coz otherwise output inserts
         "[1] 1000 1000\n", # blank lines that no longer fit on screen
         "[1] NA NA\n",
         "[1]  0.00000 35.50599")
```
```{r dodgr-dists-result2, eval = FALSE}
round (1000 ^ 2 * length (which (is.na (d))) / length (d))
```
```{r dodgr-dists-result2-fakey, eval = TRUE, echo = FALSE, message = TRUE}
message ("[1] 32709")
```

---

class: inverse, center, top

.fonthuge[
    ~~distances~~<br><br>
    flows
    ]

---

# flows

If one person travels between each of our 1,000,000 pairs of points, how many
travel on each bit of the network?

```{r, eval = FALSE}
d <- dodgr_dists (net, from = pts, to = pts)
```
```{r dodgr-flows, eval = FALSE}
flow_mat <- matrix (1, nrow = 1000, ncol = 1000)
system.time (
    f <- dodgr_flows_aggregate (net, from = pts, to = pts,  #<<
                                flows = flow_mat)           #<<
)
```
```{r, eval = FALSE, echo = FALSE}
saveRDS (f, file = "data/dodgr-flows-ms.Rds")
```

```{r dodgr-flows-fakey, eval = TRUE, echo = FALSE, message = TRUE}
message ("   user  system elapsed\n",
"145.186   1.107  20.016")
```

---

# flows

```{r dodgr-flows-result-d, eval = TRUE, message = TRUE}
head (net)
```

---

# flows

```{r dodgr-flows-result, eval = TRUE, message = TRUE}
head (f)
```

---

# flows

```{r merge-directed-flows, eval = TRUE, message = TRUE}
fm <- merge_directed_flows (f)
```

- Adds flows .bold[A] `r icon::fa_long_arrow_alt_right()` .bold[B] to .bold[B]
  `r icon::fa_long_arrow_alt_right()` .bold[A]
- Removes one of the two edges
- Also removes all edges with zero flow

--

```{r merge-directed-flows-result, message = TRUE}
dim (f); dim (fm)
nrow (fm) / nrow (f)
```

---

# flows

```{r flowplot, out.height = "400px", out.width = "500px", echo = TRUE}
library (ggplot2)
ggplot (fm, aes (flow)) + geom_freqpoly ()
```

---
class: center, inverse

.fonthuge[
mapdeck <br><br> time!
]

---

# flows

```{r dodgr-flows-result2, eval = TRUE, message = TRUE}
head (fm)
```

---

# flows

```{r ms-mapdeck, eval = FALSE}
library (mapdeck)
f$flow <- 20 * f$flow / max (f$flow)
pal <- colorRampPalette (c ("lawngreen", "red"))
loc <- apply (sf::st_coordinates (ms$osm_lines), 2, mean) [1:2]
mapdeck (style = 'mapbox://styles/mapbox/dark-v9',
         zoom = 12,
         location = loc) %>%
    add_line (data = fm,
              layer_id = "ms-highways",
              origin = c("from_lon", "from_lat"),
              destination = c("to_lon", "to_lat"),
              stroke_colour = "flow",
              stroke_width = "flow",
              palette = pal)
```
```{r ms-mapdeck-fakey, echo = FALSE}
loc <-  c (7.619786, 51.952610)
```

---
background-image: url(img/ms-random-flows.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000
class: center, middle, inverse

---
background-image: url(img/ms-promenade100.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000
class: center, middle, inverse

---
class: center, inverse

.fontBIG[
The Promenade<br>IS<br>cycling in Münster,<br>but there are no flows along most of
it!
]

---

# How to get people riding along the Promenade?

- We first need to find the edges of our `dodgr net` which correspond to the
Promenade.
- Get OSM way IDs (making sure to exclude the "Kanalpromenade")

```{r promenade-ids, eval = FALSE}
index <- grep ("promenade", ms$osm_lines$name,
               ignore.case = TRUE)
index_kp <- grep ("kanal", ms$osm_lines$name,
                  ignore.case = TRUE)
index <- index [!index %in% index_kp]
ids_prom <- ms$osm_lines$osm_id [index] %>%
    as.character ()
```

---

# How to get people riding along the Promenade?

- Presume the Promenade "feels" 10% shorter than what it actually is

```{r promenade-reweight, eval = FALSE}
shorter <- 0.1
g <- weight_streetnet (ms$osm_lines, wt_profile = "bicycle")
index <- which (g$way_id %in% ids_prom)
g$d_weighted [index] <- g$d [index] * (1 - shorter) #<<
f <- dodgr_flows_aggregate (g, from = pts, to = pts,
                            flows = flowmat)
```

---

# How to get people riding along the Promenade?

```{r}
head (net)
```

---

# How to get people riding along the Promenade?

- Presume the Promenade "feels" 10% shorter than what it actually is

```{r promenade-reweight2, eval = FALSE}
shorter <- 0.1
g <- weight_streetnet (ms$osm_lines, wt_profile = "bicycle")
index <- which (g$way_id %in% ids_prom)
g$d_weighted [index] <- g$d [index] * (1 - shorter)
f <- dodgr_flows_aggregate (g, from = pts, to = pts,
                            flows = flowmat)
```

---
background-image: url(img/ms-promenade90.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000
class: center, middle, inverse

---

# How to get people riding along the Promenade?

Presume the Promenade "feels" (0, 10, 20, & 30)% shorter than what it actually is

```{r promenade-reweight3, eval = FALSE}
g <- weight_streetnet (ms$osm_lines, wt_profile = "bicycle")
index <- which (g$way_id %in% ids)
g$d_weighted [index] <- g$d [index] * (1 - shorter)
f <- dodgr_flows_aggregate (g, from = pts, to = pts,
                            flows = flowmat)
```

---
background-image: url(img/ms-promenade100.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000
class: center, middle, inverse

---
background-image: url(img/ms-promenade90.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000
class: center, middle, inverse

---
background-image: url(img/ms-promenade80.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000
class: center, middle, inverse

---
background-image: url(img/ms-promenade70.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000
class: center, middle, inverse

---
class: center, top, inverse
.fontBIG[
The Promenade feels like it is 30% shorter than what it actually is
]

---
class: center, top, inverse
.fontBIG[
Münsteraner_Innen are willing to ride 30% further to ride along the Promenade
]

---
class: center, top, inverse
.fontBIG[
Important Lesson: Cultural specificities are important!
]

---
class: center, top, inverse
.fontBIG[
Popquiz: <br><br>Where in Münster are cyclists most exposed to air pollution?
]

---
background-image: url(img/ms-bike-flows.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000
class: center, middle, inverse

---

# Where is Münster are cyclists most exposed to air pollution?

- Air pollution comes from automotive vehicles, so need to know where they
  travel

- Generate another flow layer

--

- Implement model to simulate dispersal to adjacent ways including bike paths

---
background-image: url(img/ms-car-flows.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000
class: center, middle, inverse

---
background-image: url(img/ms-car-flows-dispersed.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000
class: center, middle, inverse

---
background-image: url(img/ms-promenade70.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000
class: center, middle, inverse

---

# Where is Münster are cyclists most exposed to air pollution?

- Air pollution comes from automotive vehicles, so need to know where they
  travel

- Generate another flow layer

- Implement model to simulate dispersal to adjacent ways including bike paths

--

- Quantify exposure as cyclist flows times pollutant concentrations

---
background-image: url(img/ms-exposure.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000
class: center, middle, inverse

---
background-image: url(img/ms-car-flows-dispersed.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000
class: center, middle, inverse

---
background-image: url(img/ms-promenade70.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000
class: center, middle, inverse

---
background-image: url(img/ms-exposure.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000
class: center, middle, inverse

--

.fonthuge[THANK<br>YOU!]
.box-bottom[May the flow be with you]

---
class: center, middle, inverse
background-image: url(img/ms-exposure.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000

.left-column[
`r icon::fa_github(size = 2, animate = "spin")` atfutures

`r icon::fa_github(size = 2, animate = "spin")` mpadge

`r icon::fa_github(size = 2, animate = "spin")` ropensci
]

.right-column[
`r icon::fa_twitter(size = 2, animate = "spin")` bikesRdata

`r icon::fa_envelope(size = 1)` .small[mark.padgham@email.com]<br><br>

`r icon::ii_ios_world_outline(size = 5)` atfutures.github.io
]

.box-bottom[
slides at <br>
[https://github.com/mpadge/ms-user-meetup-nov2018](https://github.com/mpadge/ms-user-meetup-nov2018)
]

### 2018-11-19 myxaringantheme.R

@@ -1,0 +1,95 @@
+library(xaringanthemer)
+#background_color = "#ddddbb"
+background_color = "#ccdddd"
+inverse_background_color = "#272822"
+inverse_text_color = "#d6d6d6"
+write_xaringan_theme(text_color = "#000",
+                     header_color = "#308bd2",
+                     background_color = background_color,
+                     inverse_background_color = inverse_background_color,
+                     #link_color = "rgb(249, 38, 114)",
+                     link_color = "#23cc00",
+                     text_bold_color = NA,
+                     text_slide_number_color = inverse_background_color,
+                     padding = "1em 4em 1em 4em",
+                     background_image = NA,
+                     background_size = NA,
+                     background_position = NA,
+                     code_highlight_color = "rgba(255,255,0,0.5)",
+                     code_inline_color = "#ff6900",
+                     code_inline_background_color = NA,
+                     code_inline_font_size = "1em",
+                     inverse_text_color = inverse_text_color,
+                     inverse_text_shadow = FALSE,
+                     inverse_header_color = "#50aaee",
+                     title_slide_text_color = inverse_text_color,
+                     title_slide_background_color = inverse_background_color,
+                     title_slide_background_image = NA,
+                     title_slide_background_size = NA,
+                     title_slide_background_position = NA,
+                     footnote_color = NA,
+                     footnote_font_size = "0.9em",
+                     footnote_position_bottom = "3em",
+                     left_column_subtle_color = "#777",
+                     left_column_selected_color = "#000",
+                     blockquote_left_border_color = "lightgray",
+                     table_border_color = "#666",
+                     table_row_border_color = "#ddd",
+                     table_row_even_background_color = "#eee",
+                     text_font_size = "25px",
+                     header_h1_font_size = "55px",
+                     header_h2_font_size = "45px",
+                     header_h3_font_size = "35px",
+                     text_slide_number_font_size = "0.7em",
+                     #text_font_google = NULL,
+                     text_font_google = google_font("Montserrat", "300", "300i"),
+                     #text_font_family = "'Droid Sans'",
+                     text_font_weight = "normal",
+                     #text_font_url = "https://fonts.googleapis.com/css?family=Droid+Sans:400,700,400italic",
+                     text_font_family_fallback = "'Palatino Linotype', 'Book Antiqua', Palatino, 'Microsoft YaHei', 'Songti SC'",
+                     text_font_base = "sans",
[…]

### 2018-11-19 script.R

@@ -1,0 +1,28 @@
+library (osmdata)
+if (!file.exists ("data/osm-ms-48155.Rds"))
+{
+    # 48155 sometimes returns a very small box, other times the full area, so ...
+    bb <- c (7.632622, 51.9163722, 7.7231164, 51.9906058)
+    #dat <- opq ("48155 de") %>%
+    dat <- opq (bb) %>%
+        add_osm_feature (key = "highway") %>%
+        osmdata_sf (quiet = FALSE)
+    saveRDS (dat, file = "data/osm-ms-48155.Rds")
+}
+
+if (!file.exists ("data/osm-promenade.Rds"))
+{
+    promenade <- opq ("münster de") %>%
+        add_osm_feature (key = "highway") %>%
+        add_osm_feature (key = "name", value = "Promenade", value_exact = FALSE) %>%
+        osmdata_sf (quiet = FALSE)
+    saveRDS (promenade, file = "data/osm-promenade.Rds")
+}
+
+if (!file.exists ("data/osm-ms-highways.Rds"))
+{
+    ms <- opq ("münster de") %>%
+        add_osm_feature (key = "highway") %>%
+        osmdata_sf (quiet = FALSE)
+    saveRDS (ms, file = "data/osm-ms-highways.Rds")
+}

### 2018-11-19 ms-meetup-nov18.Rmd
:wrench:Rmd

@@
 
+- The fastest and easiest method in any language for extracting OSM data
+- Only software that allows OSM data to be extracted in 3 simple lines:
+
+```{r, eval = FALSE}
+ms <- opq ("münster de") %>%
+    add_osm_feature (key = "highway") %>%
+    osmdata_sf (quiet = FALSE)
+```
+
+- Extremely flexible; can extract any form of OSM data.
+
+---
+
+# osmdata
+
 For more information, see [the website](https://github.com/ropensci/osmdata)

### 2018-11-19 docs/slides/img/script.Rmd
reinstate docs/slides

# code to make figures

```{r}
source ("../script.R")
promenade <- readRDS ("../data/osm-promenade.Rds")
ms48155 <- readRDS ("../data/osm-ms-48155.Rds")
ms <- readRDS ("../data/osm-ms-highways.Rds")
```

## mapview images

```{r}
mapview::mapview (ms48155$osm_lines)
```

```{r}
mapview::mapview (promenade$osm_lines)
```

## Mapdeck images

These are made with mapdeck which doesn't currently integrate with xaringan, so
have to be manually screenshotted.

`img/ms48155lines-mapview.png`:
```{r mapdeck1, eval = FALSE, echo = FALSE}
library (osmdata)
library (dodgr)
library (mapdeck)
loc <- as.numeric (apply (sf::st_coordinates (ms48155$osm_lines), 2, mean) [1:2])
tok <- Sys.getenv ("MAPBOX_TOKEN")
mapdeck (style = 'mapbox://styles/mapbox/dark-v9',
         location = loc,
         zoom = 12) %>%
    add_path (data = ms48155$osm_lines,
              layer_id = "ms",
              stroke_width = 1,
              stroke_colour = "#4444dd",
              stroke_opacity = 80)
```

`img/mspromenade-mapview.png`:
```{r mapdeck2, eval = FALSE, echo = FALSE}
library (dodgr)
loc <- as.numeric (apply (sf::st_coordinates (ms48155$osm_lines), 2, mean) [1:2])
tok <- Sys.getenv ("MAPBOX_TOKEN")
mapdeck (style = 'mapbox://styles/mapbox/dark-v9',
         location = loc,
         zoom = 12) %>%
    add_path (data = promenade$osm_lines,
              layer_id = "promenade",
              stroke_width = 5,
              stroke_colour = "#33cc33",
              stroke_opacity = 100)
```

`img/ms-highway-promenade-mapview.png`:
```{r mapdeck2, eval = FALSE, echo = FALSE}
library (dodgr)
loc <- as.numeric (apply (sf::st_coordinates (ms48155$osm_lines), 2, mean) [1:2])
tok <- Sys.getenv ("MAPBOX_TOKEN")
mapdeck (style = 'mapbox://styles/mapbox/dark-v9',
         location = loc,
         zoom = 12) %>%
    add_path (data = ms48155$osm_lines,
              layer_id = "ms",
              stroke_width = 1,
              stroke_colour = "#4444dd",
              stroke_opacity = 80) %>%
    add_path (data = promenade$osm_lines,
              layer_id = "promenade",
              stroke_width = 5,
              stroke_colour = "#33cc33",
              stroke_opacity = 100)
```

# ms components

```{r}
library (dodgr)
g <- weight_streetnet (ms$osm_lines, wt_profile = "bicycle")
mapdeck (style = 'mapbox://styles/mapbox/dark-v9',
         pitch = 0,
         zoom = 12,
         location = loc) %>%
    add_line (data = g,
              layer_id = "ms-highways",
              origin = c("from_lon", "from_lat"),
              destination = c("to_lon", "to_lat"),
              stroke_colour = "component",
              stroke_width = 1,
              palette = rainbow)
```

`img/ms-random-flows.png`:
```{r}
f <- readRDS ("data/dodgr-flows-ms.Rds")
f <- dodgr::merge_directed_flows (f)
library (mapdeck)
f$flow <- 20 * f$flow / max (f$flow)
pal <- colorRampPalette (c ("lawngreen", "red"))
loc <-  c (7.619786, 51.952610)
mapdeck (style = 'mapbox://styles/mapbox/dark-v9',
         pitch = 0,
         zoom = 12,
         location = loc) %>%
    add_line (data = f,
              layer_id = "ms-highways",
              origin = c("from_lon", "from_lat"),
              destination = c("to_lon", "to_lat"),
              stroke_colour = "flow",
              stroke_width = "flow",
              palette = pal)
```

`img/ms-random-flows90.png`:
```{r}
ms <- readRDS ("data/osm-ms-highways.Rds")
library (osmdata)
library (dodgr)
ms <- osm_poly2line (ms)
# get OSM way IDs for promenade
index <- grep ("promenade", ms$osm_lines$name, ignore.case = TRUE)
kp <- grep ("kanal", ms$osm_lines$name, ignore.case = TRUE)
index <- index [!index %in% kp]
ids <- ms$osm_lines$osm_id [index] %>% as.character ()
index <- which (g$way_id %in% ids)

v <- dodgr_vertices (g)
pts <- sample (v$id, size = 1000)
flowmat <- matrix (1, nrow = 1000, ncol = 1000)

g <- weight_streetnet (ms$osm_lines, wt_profile = "bicycle")
g$d_weighted [index] <- g$d [index] * 0.9
f <- dodgr_flows_aggregate (g, from = pts, to = pts, flows = flowmat)
saveRDS (f, file = "data/dodgr-flows-ms90.Rds")

g <- weight_streetnet (ms$osm_lines, wt_profile = "bicycle")
g$d_weighted [index] <- g$d [index] * 0.8
f <- dodgr_flows_aggregate (g, from = pts, to = pts, flows = flowmat)
saveRDS (f, file = "data/dodgr-flows-ms80.Rds")

g <- weight_streetnet (ms$osm_lines, wt_profile = "bicycle")
g$d_weighted [index] <- g$d [index] * 0.7
f <- dodgr_flows_aggregate (g, from = pts, to = pts, flows = flowmat)
saveRDS (f, file = "data/dodgr-flows-ms70.Rds")
```
```{r}
plot1map <- function (f)
{
    if (is.character (f)) # name of file
    {
        f <- readRDS (f)
        f <- dodgr::merge_directed_flows (f)
    }
    library (mapdeck)
    f$flow <- 20 * f$flow / max (f$flow)
    pal <- colorRampPalette (c ("lawngreen", "red"))
    loc <-  c (7.619786, 51.962)
    mapdeck (style = 'mapbox://styles/mapbox/dark-v9',
             pitch = 0,
             zoom = 14,
             location = loc) %>%
    add_line (data = f,
              layer_id = "ms-highways",
              origin = c("from_lon", "from_lat"),
              destination = c("to_lon", "to_lat"),
              stroke_colour = "flow",
              stroke_width = "flow",
              palette = pal)
}
plot1map ("data/dodgr-flows-ms.Rds")
plot1map ("data/dodgr-flows-ms90.Rds")
plot1map ("data/dodgr-flows-ms80.Rds")
plot1map ("data/dodgr-flows-ms70.Rds")
```

# exposure

## vehicular flows

Same adjustment has to be made to the Ring, which is "secondary", and to lots of
the inner-city roads which are "primary" to get flows concentrated around the
ring rather than through the middle
```{r}
library (osmdata)
ms <- readRDS ("data/osm-ms-highways.Rds")
ms <- osm_poly2line (ms)

# get OSM way IDs for Ring, which are all secondary
index_ring <- which (grepl ("ring", ms$osm_lines$name, ignore.case = TRUE) &
                ms$osm_lines$highway == "secondary")
index_sg <- which (grepl ("am stadtgraben", ms$osm_lines$name, ignore.case = TRUE) &
                   ms$osm_lines$highway == "primary")
index_sp <- which (grepl ("schlossplatz", ms$osm_lines$name, ignore.case = TRUE) &
                   ms$osm_lines$highway == "primary")
index_nt <- which (grepl ("neutor", ms$osm_lines$name, ignore.case = TRUE) &
                   ms$osm_lines$highway == "primary")
index_st <- which (grepl ("steinfurter", ms$osm_lines$name, ignore.case = TRUE) &
                   ms$osm_lines$highway == "primary")
index_s <- unique (c (index_sg, index_sp, index_nt, index_st))
ids_ring <- ms$osm_lines$osm_id [index_ring]
ids_s <- ms$osm_lines$osm_id [unique (c (index_sg, index_sp, index_nt, index_st))]
```

```{r}
weighting_profiles [weighting_profiles$name == "motorcar", ]
```
wt_profile for secondary is 0.8, and primary is 0.9

functions from `flowlayers`:
```{r}
dist_to_lonlat_range <- function (verts, d = 20)
{
    xy0 <- c (mean (verts$x), mean (verts$y))
    names (xy0) <- c ("x", "y")
    minf <- function (a, xy0) { abs (geodist::geodist (xy0, xy0 + a) - d) }
    optimise (minf, c (0, 0.1), xy0)$minimum
}

disperse_flows <- function (graph, d = 20)
{
    # This actually works by mapping flows to vertices, and then applying a
    # Gaussian kernel to those points, before re-mapping them back to edges
    v <- dodgr::dodgr_vertices (graph)
    sig <- dist_to_lonlat_range (v, d)

    indxf <- match (v$id, graph$from_id)
    indxt <- match (v$id, graph$to_id)
    indxf [is.na (indxf)] <- indxt [is.na (indxf)]
    indxt [is.na (indxt)] <- indxf [is.na (indxt)]
    v$flow <- apply (cbind (graph$flow [indxf], graph$flow [indxt]), 1, max)

    # warnings are issued here if any (x, y) are duplicated
    xy <- suppressWarnings (spatstat::ppp (v$x, v$y, range (v$x), range (v$y)))
    d <- spatstat::density.ppp (xy, weights = v$flow, sigma = sig, at = "points")
    vsum <- sum (v$flow)
    d <- d * vsum / sum (d)
    indx <- which (d > v$flow)
    v$flow [indx] <- d [indx]
    v$flow <- v$flow * vsum / sum (v$flow)

    # Then map those vertex values back onto the graph
    indxf <- match (graph$from_id, v$id)
    indxt <- match (graph$to_id, v$id)
    fmax <- apply (cbind (v$flow [indxf], v$flow [indxt]), 1, max)
    fsum <- sum (graph$flow)
    fmax <- fmax * fsum / sum (graph$flow)
    indx <- which (fmax > graph$flow)
    # That again imbalances the flow of the graph, so needs to be standardised
    # back to original sum again
    fsum <- sum (graph$flow)
    graph$flow [indx] <- fmax [indx]
    graph$flow <- graph$flow * fsum / sum (graph$flow)

    return (graph)
}
```

```{r}
library (dodgr)
gv <- weight_streetnet (ms$osm_lines, wt_profile = "motorcar")
rrr <- 4
index <- match (ids_ring, gv$way_id)
gv$d_weighted [index] <- gv$d [index] / rrr
index <- match (ids_s, gv$way_id)
gv$d_weighted [index] <- gv$d [index] * rrr
gc <- dodgr_contract_graph (gv)
v <- dodgr_vertices (gc$graph)
pts <- sample (v$id, size = 1000)
message ("aggregating ... ", appendLF = FALSE)
fv <- dodgr_flows_aggregate (gv, from = pts, to = pts,
                             flow = matrix (1, nrow = 1000, ncol = 1000))
message ("done\ndispersing ... ", appendLF = FALSE)
fvd <- disperse_flows (fv, 20) %>%
    merge_directed_flows ()
fv <- merge_directed_flows (fv)
mesage ("done")
```
The two maps of vehicular flows:
```{r}
plot1map (fv)
plot1map (fvd)
```
```{r}
f <- readRDS ("data/dodgr-flows-ms70.Rds") %>%
    merge_directed_flows ()
f <- f [which (f$edge_id %in% fvd$edge_id), ]
indx1 <- match (f$edge_id, fvd$edge_id)
indx2 <- match (fvd$edge_id [indx1], f$edge_id)
f$flow <- f$flow / max (f$flow)
fvd$flow <- fvd$flow / max (fvd$flow)
f$flow <- f$flow * fvd$flow [indx1] [indx2]
plot1map (f)
```

### 2018-11-20 img/script.Rmd
update scripts

--- 2018-11-19 docs/slides/img/script.Rmd
@@
 loc <- as.numeric (apply (sf::st_coordinates (ms48155$osm_lines), 2, mean) [1:2])
-tok <- Sys.getenv ("MAPBOX_TOKEN")
 mapdeck (style = 'mapbox://styles/mapbox/dark-v9',
@@
               stroke_opacity = 80)
+```
+
+### timing comparison mapview vs mapdeck
+
+```{r}
+library (mapview)
+library (mapdeck)
+system.time (mapview (ms48155$osm_lines))
+loc <- as.numeric (apply (sf::st_coordinates (ms48155$osm_lines), 2, mean) [1:2])
+system.time (
+mapdeck (style = 'mapbox://styles/mapbox/dark-v9',
+         location = loc,
+         zoom = 12) %>%
+    add_path (data = ms48155$osm_lines,
+              layer_id = "ms",
+              stroke_width = 1,
+              stroke_colour = "#4444dd",
+              stroke_opacity = 80)
+)
 ```
@@
 fv <- merge_directed_flows (fv)
-mesage ("done")
+message ("done")
 ```
@@
 f$flow <- f$flow * fvd$flow [indx1] [indx2]
+saveRDS (f, file = "dodgr-exposure-ms.Rds")
 plot1map (f)

### 2018-11-20 ms-meetup-nov18.Rmd
update scripts

@@
 source ("myxaringantheme.R")
-source ("script.R")
+source ("data-script.R")
 promenade <- readRDS ("data/osm-promenade.Rds")
@@
 - `opq` `r icon::fa_long_arrow_alt_right()` .bold[WHERE]
-- `opq` `r icon::fa_long_arrow_alt_right()` .bold[WHERE]
 - `add_osm_feature` `r icon::fa_long_arrow_alt_right()` .bold[WHAT]
@@
     add_osm_feature (key = "highway") %>%
-    add_osm_feature (key = "name", value = "Promenade",
+    add_osm_feature (key = "name", value = "promenade",
                      value_exact = FALSE) %>%
@@
 
-```{r prom-sf, message = TRUE}
-promenade
-```
-
----
-
-# Visualising osmdata
-
-```{r prom-sf-print}
-library (sf)
-promenade$osm_lines
-```
----
-
-# Visualising osmdata
-
 ```{r, eval = FALSE}
@@
 
+```{r, eval = FALSE}
+system.time (
+             mapview (ms48155$osm_lines)
+)
+```
+```{r, echo = FALSE, message = TRUE}
+res <- c (3.799, 0.074, 3.906)
+names (res) <- c ("user", "system", "elapsed")
+res
+```
+
+```{r, eval = FALSE}
+system.time (
+mapdeck (style = 'mapbox://styles/mapbox/dark-v9',
+         location = loc,
+         zoom = 12) %>%
+    add_path (data = ms48155$osm_lines,
+              layer_id = "ms",
+              stroke_width = 1,
+              stroke_colour = "#4444dd",
+              stroke_opacity = 80)
+)
+```
+```{r, echo = FALSE, message = TRUE}
+res <- c (0.222, 0.000, 0.237)
+names (res) <- c ("user", "system", "elapsed")
+res
+```
+
+---
+
+# Visualising osmdata
+
 One further alternative for static plots is the [`osmplotr`
@@
 ```{r, eval = FALSE}
+ms <- opq ("münster de") %>%               # WHERE
+    add_osm_feature (key = "highway") %>%  # WHAT
+    osmdata_sf (quiet = FALSE)             # HOW
+```
+
+- Extremely flexible; can extract any form of OSM data.
+
+```{r, eval = FALSE}
 ms <- opq ("münster de") %>%
-    add_osm_feature (key = "highway") %>%
+    add_osm_feature (key = "building") %>%
     osmdata_sf (quiet = FALSE)
 ```
-
-- Extremely flexible; can extract any form of OSM data.
-
----
-
-# osmdata
-
-For more information, see [the website](https://github.com/ropensci/osmdata)
+- For more information, see [the website](https://github.com/ropensci/osmdata)
 
@@
 
-```{r weight-streetnet-ms2, eval = FALSE, message = TRUE}
-ms <- osm_poly2line (ms) # osmdata function
-library (dodgr)
-net <- weight_streetnet (ms$osm_lines, wt_profile = "bicycle") #<<
-```
-
----
-
-# dodgr
-
-Convert Münster OSM data to `dodgr` format
-
 ```{r weight-streetnet-ms3, eval = FALSE, message = TRUE}
@@
 
-```{r dodgr-vertices1-fakey, eval = FALSE}
-v <- dodgr_vertices (net) #<<
-dim (v); head (v)
-```
-```{r dodgr-vertices1, eval = TRUE, echo = FALSE}
-v <- dodgr_vertices (net)
-dim (v); head (v, n = 5)
-```
-
-- `id` is the unique OSM is of each vertex<br>("node" in OSM terminology)
-- most `dodgr` functions work either with<br>coordinates or `id` values
-
----
-
-# dodgr distances
-
-Calculate cycling distances between random points.
-
 ```{r dodgr-vertices2, eval = FALSE}
@@
 
-```{r dodgr-dists1, eval = FALSE}
+```{r dodgr-dists1-copy, eval = FALSE}
 system.time (
@@
 ```
-```{r dodgr-dists1-fakey, echo = FALSE, message = TRUE}
-message ("   user  system elapsed\n",
-" 43.847   0.013   6.042")
-```
-
---
-
-- Those 1,000,000 distances could have been calculated with the google API
-- And that would cost `1e6 * 0.004 = $US4,000`
-- $US 4,000 for 6 seconds of calculation!
-- To date, `dodgr` provides one of the only .bold[free] ways to perform such
-  calcualtions, and is also entirely .bold[open source].
-
----
-
-# dodgr distances
-
-```{r dodgr-dists1-copy, eval = FALSE}
-system.time (
-             d <- dodgr_dists (net, from = pts, to = pts)
-             )
-```
 ```{r dodgr-dists1-copy-fakey, echo = FALSE, message = TRUE}
@@
 ```
+
+--
+
 ```{r dodgr-dists-result1, eval = FALSE}
@@
 ```
+
+---
+
+# dodgr distances
+
+```{r dodgr-dists1, eval = FALSE}
+system.time (
+             d <- dodgr_dists (net, from = pts, to = pts) #<<
+             )
+```
+```{r dodgr-dists1-fakey, echo = FALSE, message = TRUE}
+message ("   user  system elapsed\n",
+" 43.847   0.013   6.042")
+```
+
+- Those 1,000,000 distances could have been calculated with the google API
+- And that would cost `1e6 * 0.004 = $US4,000`
+- $US 4,000 for 6 seconds of calculation!
+- To date, `dodgr` provides one of the only .bold[free] ways to perform such
+  calculations, and is also entirely .bold[open source].
+
+---
+class: left, top, inverse
+background-image: url(img/ms-promenade.png)
+background-size: contain 
+background-position: 50% 50%
+background-color: #000000
 
@@
 
+```{r dodgr-flows-result2, eval = TRUE, message = TRUE}
+head (fm)
+```
+
+---
+
+# flows
+
 ```{r flowplot, out.height = "400px", out.width = "500px", echo = TRUE}
@@
 .fonthuge[
-mapdeck <br><br> time!
-]
-
----
-
-# flows
-
-```{r dodgr-flows-result2, eval = TRUE, message = TRUE}
-head (fm)
-```
-
+mapdeck <br> time!<br>
+`r icon::fa_uber(animate = "spin")`
+]
 ---
@@
 
---
-
 - Quantify exposure as cyclist flows times pollutant concentrations

### 2018-11-20 data-script.R

@@ -1,0 +1,186 @@
+library (osmdata)
+
+if (!dir.exists ("data"))
+    dir.create ("data")
+
+if (!file.exists ("data/osm-ms-48155.Rds"))
+{
+    # 48155 sometimes returns a very small box, other times the full area, so ...
+    bb <- c (7.632622, 51.9163722, 7.7231164, 51.9906058)
+    #dat <- opq ("48155 de") %>%
+    dat <- opq (bb) %>%
+        add_osm_feature (key = "highway") %>%
+        osmdata_sf (quiet = FALSE)
+    saveRDS (dat, file = "data/osm-ms-48155.Rds")
+}
+
+if (!file.exists ("data/osm-promenade.Rds"))
+{
+    promenade <- opq ("münster de") %>%
+        add_osm_feature (key = "highway") %>%
+        add_osm_feature (key = "name", value = "Promenade", value_exact = FALSE) %>%
+        osmdata_sf (quiet = FALSE)
+    saveRDS (promenade, file = "data/osm-promenade.Rds")
+}
+
+if (!file.exists ("data/osm-ms-highways.Rds"))
+{
+    ms <- opq ("münster de") %>%
+        add_osm_feature (key = "highway") %>%
+        osmdata_sf (quiet = FALSE)
+    saveRDS (ms, file = "data/osm-ms-highways.Rds")
+}
+
+if (!(file.exists ("data/dodgr-flows-ms.Rds") &
+      file.exists ("data/dodgr-flows-ms90.Rds") &
+      file.exists ("data/dodgr-flows-ms80.Rds") &
+      file.exists ("data/dodgr-flows-ms70.Rds")))
+{
+    ms <- readRDS ("data/osm-ms-highways.Rds")
+    library (dodgr)
+    ms <- osm_poly2line (ms)
+    # get OSM way IDs for promenade
+    index <- grep ("promenade", ms$osm_lines$name, ignore.case = TRUE)
+    kp <- grep ("kanal", ms$osm_lines$name, ignore.case = TRUE)
+    index <- index [!index %in% kp]
+    ids <- ms$osm_lines$osm_id [index] %>% as.character ()
+    index <- which (g$way_id %in% ids)
+
+    g0 <- weight_streetnet (ms$osm_lines, wt_profile = "bicycle")
+    v <- dodgr_vertices (g)
+    pts <- sample (v$id, size = 1000)
[…]

### 2018-11-20 script.R

@@ -1,28 +1,0 @@
-library (osmdata)
-if (!file.exists ("data/osm-ms-48155.Rds"))
-{
-    # 48155 sometimes returns a very small box, other times the full area, so ...
-    bb <- c (7.632622, 51.9163722, 7.7231164, 51.9906058)
-    #dat <- opq ("48155 de") %>%
-    dat <- opq (bb) %>%
-        add_osm_feature (key = "highway") %>%
-        osmdata_sf (quiet = FALSE)
-    saveRDS (dat, file = "data/osm-ms-48155.Rds")
-}
-
-if (!file.exists ("data/osm-promenade.Rds"))
-{
-    promenade <- opq ("münster de") %>%
-        add_osm_feature (key = "highway") %>%
-        add_osm_feature (key = "name", value = "Promenade", value_exact = FALSE) %>%
-        osmdata_sf (quiet = FALSE)
-    saveRDS (promenade, file = "data/osm-promenade.Rds")
-}
-
-if (!file.exists ("data/osm-ms-highways.Rds"))
-{
-    ms <- opq ("münster de") %>%
-        add_osm_feature (key = "highway") %>%
-        osmdata_sf (quiet = FALSE)
-    saveRDS (ms, file = "data/osm-ms-highways.Rds")
-}

### 2018-11-20 setup-script.R

@@ -1,0 +1,30 @@
+# Script to open mapdeck tabs for actual presentation
+library (mapdeck)
+set_token (Sys.getenv ("MAPBOX_TOKEN"))
+
+plot1map <- function (f)
+{
+    if (is.character (f)) # name of file
+    {
+        f <- readRDS (f)
+        f <- dodgr::merge_directed_flows (f)
+    }
+    library (mapdeck)
+    f$flow <- 20 * f$flow / max (f$flow)
+    pal <- colorRampPalette (c ("lawngreen", "red"))
+    loc <-  c (7.619786, 51.962)
+    mapdeck (style = 'mapbox://styles/mapbox/dark-v9',
+             pitch = 0,
+             zoom = 14,
+             location = loc) %>%
+    add_line (data = f,
+              layer_id = "ms-highways",
+              origin = c("from_lon", "from_lat"),
+              destination = c("to_lon", "to_lat"),
+              stroke_colour = "flow",
+              stroke_width = "flow",
+              palette = pal)
+}
+
+plot1map ("data/dodgr-flows-ms70.Rds")
+plot1map ("data/dodgr-exposure-ms.Rds")

### 2018-11-20 ms-meetup-nov18.Rmd
:rocket:

@@
 
-```{r osmdata-sp, eval = FALSE}
-library (osmdata)
-opq ("48155 de") %>%
-    add_osm_feature (key = "highway") %>%
-    osmdata_sp ()
-```
-
-- `osmdata_sf` `r icon::fa_long_arrow_alt_right()` .bold[Simple Features]
-- `osmdata_sp` `r icon::fa_long_arrow_alt_right()` .bold[R Spatial]
-
----
-
-# The osmdata package - Spatial Formats
-
-```{r osmdata-sc, eval = FALSE}
-library (osmdata)
-opq ("48155 de") %>%
-    add_osm_feature (key = "highway") %>%
-    osmdata_sc ()
-```
-
-- `osmdata_sf` `r icon::fa_long_arrow_alt_right()` .bold[Simple Features]
-- `osmdata_sp` `r icon::fa_long_arrow_alt_right()` .bold[R Spatial]
-- `osmdata_sc` `r icon::fa_long_arrow_alt_right()` .bold[Silicate (Simplicial Complex)]
-
----
-
-# The osmdata package - Spatial Formats
-
-```{r osmdata-xml, eval = FALSE}
-library (osmdata)
-opq ("48155 de") %>%
-    add_osm_feature (key = "highway") %>%
-    osmdata_xml (file = "ms48155.osm")
-```
-
-- `osmdata_sf` `r icon::fa_long_arrow_alt_right()` .bold[Simple Features]
-- `osmdata_sp` `r icon::fa_long_arrow_alt_right()` .bold[R Spatial]
-- `osmdata_sc` `r icon::fa_long_arrow_alt_right()` .bold[Silicate (Simplicial Complex)]
-- `osmdata_xml` `r icon::fa_long_arrow_alt_right()` .bold[XML / OSM]
-
----
-
-# The osmdata package - Spatial Formats
-
 ```{r osmdata-pbf, eval = FALSE}
@@
     add_osm_feature (key = "highway") %>%
-    osmdata_pbf (file = "ms48155.pbf")
+    osmdata_sf (file = "ms48155.pbf")
 ```
@@
 ```{r dodgr-vertices2-fakey, eval = TRUE, echo = FALSE, message = TRUE}
+v <- dodgr_vertices (net)
 dim (v); head (v, n = 5)
@@
 
+--
+
+.fontBIG[
+And can be incorporated!
+]
+
 ---

### 2018-11-20 setup-script.R

@@ -26,5 +26,5 @@ plot1map <- function (f)
               palette = pal)
 }
 
-plot1map ("data/dodgr-flows-ms70.Rds")
+plot1map ("data/dodgr-flows-ms.Rds")
 plot1map ("data/dodgr-exposure-ms.Rds")

### 2019-01-14 ms-meetup-nov18.Rmd
new deploy
final deploy
fix example code to visualize the *P*romenade

@@
 
-```{r osmdata-pbf, eval = FALSE}
+```{r osmdata-all, eval = FALSE}
 library (osmdata)
@@
     add_osm_feature (key = "highway") %>%
-    osmdata_sf (file = "ms48155.pbf")
+    osmdata_sf ()
 ```
@@
     add_osm_feature (key = "highway") %>%
-    add_osm_feature (key = "name", value = "promenade", #<<
+    add_osm_feature (key = "name", value = "Promenade", #<<
                      value_exact = FALSE) %>%           #<<
@@
     add_osm_feature (key = "highway") %>%
-    add_osm_feature (key = "name", value = "promenade",
+    add_osm_feature (key = "name", value = "Promenade",
                      value_exact = FALSE) %>%
@@
 
-# dodgr distances
+# dodgr
 

### 2019-01-14 ms-meetup-nov18.Rmd
small fixes to have working code on the slides

@@
 library (mapdeck)
-f$flow <- 20 * f$flow / max (f$flow)
+fm$flow <- 20 * fm$flow / max (fm$flow)
 pal <- colorRampPalette (c ("lawngreen", "red"))
-loc <- apply (sf::st_coordinates (ms$osm_lines), 2, mean) [1:2]
+loc <- as.numeric(apply (sf::st_coordinates (ms$osm_lines), 2, mean)) [1:2]
 mapdeck (style = 'mapbox://styles/mapbox/dark-v9',
@@
 g$d_weighted [index] <- g$d [index] * (1 - shorter)
+flow_mat <- matrix (1, nrow = 1000, ncol = 1000)
 f <- dodgr_flows_aggregate (g, from = pts, to = pts,
-                            flows = flowmat)
+                            flows = flow_mat)
 ```
@@
 
-# Where is Münster are cyclists most exposed to air pollution?
+# Where in Münster are cyclists most exposed to air pollution?
 

### 2019-01-14 ms-meetup-nov18.Rmd
no vertical scroll on slides

@@
 pal <- colorRampPalette (c ("lawngreen", "red"))
-loc <- as.numeric(apply (sf::st_coordinates (ms$osm_lines), 2, mean)) [1:2]
+loc <- apply (sf::st_coordinates (ms$osm_lines), 2, mean) [1:2]
 mapdeck (style = 'mapbox://styles/mapbox/dark-v9',
          zoom = 12,
-         location = loc) %>%
+         location = as.numeric(loc)) %>%
     add_line (data = fm,
@@
 g$d_weighted [index] <- g$d [index] * (1 - shorter) #<<
+flow_mat <- matrix (1, nrow = 1000, ncol = 1000)
 f <- dodgr_flows_aggregate (g, from = pts, to = pts,
-                            flows = flowmat)
+                            flows = flow_mat)
 ```
